# Presentation of data

## Using `knitr::kable()`

Here is our trial data, presented using `knitr::kable()`.

```{r data_presentation_kable}
kable( trial_data.clean )
```

There are some things that we can't do using kable.
This includes:

 * merging cells in the header
 * conditional formatting (font colour, fill, style etc.)
 * dynamic formatting (i.e., changes with value in a variable)
 * white space within cells (useful for subtypes) - not covered here

## Using `flextable()`

First, some basics.

1. formatting cell contents
2. displaying a preview in Word
3. autofitting the contents
4. changing header titles

These can all be done by the more basic `kable()` function.

```{r data_presentation_flextable_1}

trial_data.flextable = flextable( trial_data.clean ) %>%
  ### Formatting datatypes consistently
  set_formatter_type( fmt_date    = "%d-%m-%Y",
                      fmt_double  = "%.02f",
                      fmt_integer = "%.0f",
                      na_str      = "-" ) %>%
  ### Autofit contents
  autofit() %>%

  ### Formatting column headings
  set_header_labels(values=c(
    "patient"            = "Patient",
    "date_of_treatment"  = "Treatment date",
    "drug_treatment"     = "Treatment",
    "primary_endpoint_1" = "(1)",
    "primary_endpoint_2" = "(2)"
  ))

print( trial_data.flextable, preview="docx" )

```

Here are a few thing that we can do with flextable (that you can't do
with `kable()`...or at least I couldn't work out how to do them
with `kable()`).

5. Merging cells in the header
6. Fine control over borders

```{r data_presentation_flextable_2}

trial_data.flextable =  trial_data.flextable %>%
  
  ### Merging cells
  add_header_row( values   =c("","","","Primary endpoints"),
                  colwidths=c( 1, 1, 1,                  2) ) %>%
  
  ### Formatting cells (borders, background)
  bg( part="header", bg="grey" ) %>%
  bold( bold = TRUE, part = "header" ) %>%
  align(align = "center", part="header", i=1) %>%
  border_remove() %>%
  hline_top(part="header",border=tbl_border.heavy) %>%
  hline_top( part="body", border=tbl_border.heavy) %>%
  hline_bottom( part="body", border=tbl_border.heavy) %>%
  border(i=1,j=4:5,part="header", border.bottom = tbl_border.standard)

print( trial_data.flextable, preview="docx" )

```

Something that I find really useful is the conditional formatting.
This lets you emphasize particular rows/columns/cells to which you
want to draw the reader's attention.

```{r data_presentation_conditional_formatting}

highlight_colour.yellow = "#F9E79F"
highlight_colour.red    = "#F5B7B1"

trial_data.flextable %>%
  #bg( i=~primary_endpoint_1<0, bg=highlight_colour.yellow ) %>%
  #bg( i=~primary_endpoint_2>5, bg=highlight_colour.red    ) #%>%
  #bg( i = ~ primary_endpoint_1 < 0 & primary_endpoint_2 > 5,
 #     bg=highlight_colour.yellow ) #%>%
 bg( i = ~ primary_endpoint_1 < 0 & primary_endpoint_2 > 5,
    # j = ~ primary_endpoint_1 + primary_endpoint_2,
     bg=highlight_colour.yellow ) #%>%
  # color(i = ~ is.na(primary_endpoint_1) & is.na(primary_endpoint_2), 
        # color="grey")

```

Don't have to use hardcoded values - you can use values that you
have stored in variables.

Here I am highlighting those cells that are greater than the mean
for each of the endpoints.

```{r data_presentation_dynamic_formatting}

highlight_colour.blue = "#85C1E9"

primary_endpoint_1.mean = mean( trial_data.clean$primary_endpoint_1, na.rm=TRUE )
primary_endpoint_2.mean = mean( trial_data.clean$primary_endpoint_2, na.rm=TRUE )

trial_data.flextable %>%
  
  bg( i = ~ primary_endpoint_1 > primary_endpoint_1.mean,
      j = ~ primary_endpoint_1,
      bg=highlight_colour.blue ) %>%
  
  bg( i = ~ primary_endpoint_2 > primary_endpoint_2.mean,
      j = ~ primary_endpoint_2,
      bg=highlight_colour.blue ) %>%
  
  color(i = ~ is.na(primary_endpoint_1) & is.na(primary_endpoint_2), 
        color="grey")

```
